White Dwarf Mergers on Adaptive Meshes\\ II. Nuclear Reactions

### 2017-09-22 papers/wdmerger_II/paper.tex
NERSC and OLCF are now officially recognized facilities
Update paper II setup description

We implement the white dwarf collision problem in \castro\ in the \wdmerger\ problem
directory. We prepare the domain in the way most previous papers have done so:
the white dwarf centers of mass are initially separated by a distance of four times
the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
consequently the initial distance, depends on the equation of state used, and this
can vary somewhat depending on the relevant physics used, such as Coulomb corrections
in the Helmholtz equation of state. However, the results do not really depend on the
exact initial distance as long as there is enough time for the WDs to distort in
response to tidal forces as they approach. Their initial velocity is that of
two point-masses in free-fall towards each other coming in from infinity, such that
the contact point is at the origin, and they approach each other along a coordinate axis.
In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
composed of equal parts carbon and oxygen by mass, for comparison to earlier studies;
however, the code does have the ability to vary the impact parameter of the collision
and the mass and composition of the WDs.

@@

As in Paper I, the region outside the WDs is filled with a low-density ambient
material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition is the
same as that of the WDs. Reactions in this region are unimportant, so for
computational efficiency, we disable all nuclear burning for zones that have
$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
the results do not depend on this choice. The initial ambient
temperature is $10^7\ \text{K}$, which is the same as the initial temperature
everywhere throughout the WDs, and we set the temperature floor to the same
value (any zone that reaches below this temperature will be reset). Setting the
temperature floor to a lower value has only minor effects on the simulation outcome.

Our primary metric for this test is the amount of $^{56}$Ni generated in the collision,
as this is the parameter most directly related to the observable quantities of interest
for Type Ia supernovae. We collect information at the end of every timestep about the
total amount of nickel on the grid (in solar masses), and we use the maximum value of
this nickel mass.

@@

>>>>

We implement the white dwarf collision problem in \castro\ in the \wdmerger\ problem
directory. We prepare the domain in the way most previous papers have done so:
the white dwarf centers of mass are initially separated by a distance of four times
the (secondary) white dwarf radius. Note that the radius of the white dwarf, and
consequently the initial distance, depends on the equation of state used, and this
can vary somewhat depending on the relevant physics used, such as Coulomb corrections
in the Helmholtz equation of state. However, the results do not really depend on the
exact initial distance as long as there is enough time for the WDs to distort in
response to tidal forces as they approach. Their initial velocity is that of
two point-masses in free-fall towards each other coming in from infinity, such that
the contact point is at the origin, and they approach each other along a coordinate axis.
In this paper we focus only on head-on collisions of two 0.64 $\msolar$ white dwarfs
composed of equal parts carbon and oxygen by mass, at an initially uniform temperature
of $T = 10^7$ K. This choice is for comparison to earlier studies; the code does have
the ability to vary the impact parameter of the collision and the mass and composition
of the WDs.

@@

As in Paper I, the region outside the WDs is filled with a low-density ambient
material with $\rho = 10^{-4}\ \text{g / cm}^{-3}$ whose composition and temperature is the
same as that of the WDs. Reactions in this region are unimportant, so for
computational efficiency, we disable all nuclear burning for zones that have
$\rho < 10^6\ \text{g / cm}^{-3}$ and $T < 10^8\ \text{K}$. We have confirmed that
the results do not depend on this choice. The density floor for the simulation is set
at $10^{-5}\ \text{g / cm}^{-3}$ and the temperature floor is $10^7$ K.

Our primary metric for this test is the amount of $^{56}$Ni generated in the collision,
as this is the parameter most directly related to the observable quantities of interest
for Type Ia supernovae. We collect information at the end of every timestep about the
total amount of nickel on the grid (in solar masses), and we use the maximum value of
this nickel mass. Plotfiles of a few relevant quantities (density, temperature, velocity,
pressure, sound speed, and the ratio $\tau_s / \tau_e$) are created every 0.01 seconds.
The timesteps are limited to ensure that every multiple of this plotfile interval is
reached exactly, which implies an effective maximum timestep of 0.01 seconds.

@@

\facilities{OLCF, NERSC}
